### 2015-04-22 thesis/thesis/chap3/chap3.tex
changes

\chapter{Probabilistic Social Commitments}\label{cha:PCTLC}

In this chapter\footnote{The results of this chapter have been published in the journal of Applied Soft Computing \cite{Sultan2014a}, and in SoMet\_13 \cite{Sultan2013}.}, we establish a formal approach that allows us to precisely address probabilistic social commitments in MASs. The proposed approach is based on a new logic called the Probabilistic Logic of Commitments (PCTLC). This logic is intended to be used for specifying, reasoning about, and verifying social commitments in the presence of uncertainty.
PCTLC extends PCTL \cite{Hansson1994} with a commitment modality. We model MASs using a new version of interpreted systems that merges two extended versions of the original formalism namely, the probabilistic interpreted systems \cite{Halpern2003,Wan2013}, and the communicative interpreted systems \cite{Bentahar2012,El-Menshawy2013a}. Finally, we propose a model checking technique for verifying the new logic (PCTLC). %social commitments employed in uncertain settings.

\section{Introduction} \label{sec:introduction}

In order to represent and reason about social commitments in MASs,
commitment logics that extend CTL (Computation Tree Logic), LTL
(Linear Temporal Logic), and CTL$^{*}$ (superset of CTL and LTL),
have been proposed, see for example
\cite{Bentahar2004,Giordano2007,Pham2007,Singh2000,Verdicchio2003}.

However, current logics are merely related to specifying and
verifying social commitments under the assumption of reliable behavior. That is, they assume an absolute, non-probabilistic running of systems under consideration.
Unfortunately, this is not always the case. Heterogeneous and
autonomous intelligent components in agent societies make it
challenging to precisely analyze random or unreliable agent
behaviors. This is because agents' actions are based on observing the
environment changes and in many situations agents cannot observe
all changes in the environment. Instead, each agent can only have a
partial view of other agents' behaviors \cite{LeeL09}. In such
cases, agents make estimations about the observable world as part
of their autonomous decision making processes. Moreover, when the
system being modeled is an open system, i.e., interacts with an
environment, then uncertainty in transitions may arise due to
imperfect information about the environment \cite{Walley1996}. Consequently, the problem of representing and verifying social commitments is made more complicated by the presence of transition uncertainty which makes agents uncertain about the effects of their actions on their peers and not fully
aware of the situations other agents are encountering. Moreover,
from the communication perspective, commitments themselves are
likely to be subject to probabilistic events. Xuan and Lesser
\cite{Xuan1999} have highlighted some sources of uncertainty that
make a commitment between two agents probabilistic:

\begin{enumerate}
\item The first source of uncertainty is related to the debtor's action(s). That is, debtor's action(s) might not always lead to the fulfilment of the commitment.

\item The second source comes from the agent decision processes. Debtors beliefs and desires might change such that continuing to pursue fulfilling the commitment for others becomes irrational. Debtors' beliefs about the commitment context include, for example, the degree that the agent to whom the commitment was made is still relying on its fulfillment. To the creditor, this can cause problems because its action(s) may depend on the honoring of the commitment by the debtor.

\item The third form of uncertainty comes from the incomplete knowledge of the debtor about the creditor or about the environment within which the agent interacts.
\end{enumerate}

Consequently, one cannot assume that all autonomous agents will behave as expected, and thus commitments among communicating parties cannot be treated under the assumption of certainty. Modeling uncertainty can be achieved using different tools including fuzzy logic as in \cite{Correa2013,Karray2004}, and probabilities as in \cite{Feng2011,Karthikeyan2013,Perko2011}. On the one hand, fuzzy logic is specifically designed to deal with imprecision of facts (or the membership in vaguely defined sets). Its use in MASs has been investigated by some researchers. In \cite{Garza2011a}, the authors exploited fuzzy logic in designing intelligent agents that communicate with each other using a mental approach that uses KQML \cite{Finin1994} as the underlying communication language.
However, mental approaches suffer from the semantics verification problem \cite{Wooldridge2000}. That is, they cannot verify whether an agent is acting according to a given semantics or not \cite{Wooldridge2009}. In our work, we adopt social approaches that are based on observable social commitments.

On the other hand, probability is an important component in the design and analysis of complex systems across a broad spectrum of application domains, including communication and multimedia protocols, randomised distributed algorithms, security protocols, and dynamic power management. It is commonly  used to model unreliable or unpredictable behavior. Probability deals with the chance of happening for an event or a condition (i.e., likelihood of some event or condition). Although probability has proven to be a powerful technique in handling different aspects of MASs \cite{Kulkarni2010,Wan2013}, its value in addressing social commitments for agent communication is yet to be investigated. In our research, we use probabilities to model the uncertainty because we are concerned with the likelihood of the fulfilment of the commitment. That is, when a social commitment between two agents takes place, we are interested to know about the chance of fulfilling that commitment at a certain state in the system. Additionally, using probabilities to handle the uncertainty of social commitments provides us with the privilege of exploiting existing probabilistic logics and model checkers. However, it is worth mentioning that probability assignments are not the focus of this research. Probability values can be obtained from historical data using some techniques such as the algorithm proposed in \cite{Ormandjieva2008} which allows us to compute the fixed probabilities between two states based on some other probabilities.

To motivate our study of modeling and verifying social commitments in the face of uncertainty, we use two situational examples that arise in practical settings such as web-based systems and mobile applications.

\textit{Example 1.} Let us consider the Online Shopping System
\cite{Gomaa2011} which aims at providing services for purchasing
online items. In the web-based Online Shopping System, customers
can request to purchase one or more items from the supplier.
Having selected an item, the customer commits towards the supplier
to pay in order for the request to take place. Once the order is
paid, the supplier confirms the order, and commits to deliver the
requested item and enters a planned shipping date. Finally, when
the order is shipped, the customer is notified. Because of the
uncertainty associated to the underlying infrastructures of both
commitments (i.e., the internet through which the payment is made
and the transport system used for the delivery of purchased
goods), there is no guarantee that these commitments will be
fulfilled. Therefore, reasoning about and verifying the
commitments to pay and to deliver have to be tackled with
probability in mind so that the degree of fulfilling each
commitment can be measured.

\textit{Example 2.} In the field of mobile applications which are
complex in nature, addressing social commitments should be
paired with the consideration of uncertainty of transitions and
commitments. Let us consider a simple scenario where a receiver
agent and a sender agent have an agreement, in which the receiver
agrees to pay the sender in return of the delivery of a requested
service. This can be represented as a social commitment, in which
the receiver will be committed to the sender to pay upon obtaining
the requested service. In such a scenario, because of the presence
of stochastic behavior in mobile applications, the commitment to
pay is not going to be surely satisfied.

The scenarios described above cannot be represented by existing
conventional commitment logics because of the uncertainty aspect
in both systems. Consequently, they cannot be verified. To cope
with the situation, we need a probabilistic commitment logic that
accounts for uncertainty, and a probabilistic model
checking procedure to verify properties expressed in the new logic.

The ultimate objective of this chapter is to introduce a logical approach that is capable of addressing probabilistic social commitments in MASs. This is done as follows. \underline{First}, we present a new probabilistic logic called PCTLC to express and reason about social commitments when uncertainty is a key factor. The introduction of PCTLC logic was driven by the fact that current probabilistic temporal logics such as PCTL \cite{Hansson1994} and PCTL$^*$ \cite{Baier1998} consider neither commitments nor agent communication. PCTLC extends PCTL with modalities for commitments and their fulfillments. We model probabilistic MASs by a formalism resulted from extending the  interpreted systems introduced by Fagin et al. \cite{Fagin1995}.
This extension considers agents uncertainty and their
communication abilities. Properties to be verified (i.e., social
commitments) are specified using the probabilistic logic of
commitment PCTLC. \underline{Second}, we introduce a formal and automatic, probabilistic model checking technique for probabilistic commitment-based agent interactions. Our proposed verification method is a reduction-based model checking technique and consists of transforming the problem of model
checking PCTLC into the problem of model checking PCTL \cite{Hansson1994} so that the use of PRISM is made possible. This reduction encompasses two main steps. In the first step, we devise a set of formal rules to transform the PCTLC model into an MDP model. Then, we reduce the MDP into DTMC to be as an input to the PRISM model checker. In the latter step, we transform PCTLC formulae into PCTL formulae based on some rules developed specifically for this purpose.
As argued in \cite{El-Menshawy2013a}, the main advantage of the reduction techniques compared to the direct ones is that they are easy to implement and allow the re-use of the existing model checkers. \underline{Third}, we implement the proposed model checking approach on top of the PRISM model checker and then apply it on a concrete case study, namely Oblivious Transfer Protocol \cite{Rivest1999} from cryptography domain. Figure \ref{approach-cha3} gives an overview of the proposed approach.

\begin{figure}%[h]
\begin{center}
\includegraphics [width=14cm, height=5cm]{chap3/img/approach-cha3.eps}
  \caption{A schematic view of the probabilistic social commitment approach}
\label{approach-cha3}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Probabilistic Logic of Commitments (PCTLC)}\label{sec:PCTLC}

In this section, we present a new modal logic called the probabilistic logic of commitments (PCTLC) to address probabilistic social commitments in MASs. The logic we introduce extends the probabilistic branching-time logic PCTL \cite{Hansson1994} with a commitment modality. To do so, we merge two existing logic namely PCTL \cite{Hansson1994} and CTLC \cite{Bentahar2012,El-Menshawy2013a} using the independent join technique \cite{Franceschet2004}. The independent join combines logics as they are defined in the literature. Thus, it ensures the preservation of each logic's properties in the new logic. Hereafter, we first present the syntax of our new logic, and then we define its semantics. Before going further, we define a new version of interpreted systems formalism with abilities to account for the uncertainty aspect in target MASs and to model the social interactions between communicating parties.

The PCTLC model is generated based on two extensions of the interpreted systems formalism \cite{Fagin1995} introduced in \cite{Bentahar2012,El-Menshawy2013a}, and \cite{Halpern2003,Wan2013} as discussed in Chapter \ref{cha:background}.

\begin{definition}[Models]\label{def:models}
Given a set of atomic propositions $AP = (p,q,r, \ldots)$, the
model $\mathfrak{M_1}=(S,I,\texttt{P},\{\sim_{i \rightarrow
j}\}_{{(i,j)}\in \texttt{Agt}^2},\nu)$ is a tuple where:
%
\begin{itemize}
\item  $S \subseteq L_1 \times \ldots \times L_n$ is a countable
set of all reachable global states for the system. A state $s$ is
reachable iff there exists a sequence of transitions from an
initial state to $s$ in which the probability of each transition
is greater than $0$.

\item  $I \in S$ is an initial global state for the system.

\item  $\texttt{P}:S\times S\rightarrow [0,1]$ is a total
transition probability function defined as $\texttt{P}(s,
s')=\tau(s,a^{s \rightarrow s'}, s')$ iff there exists a joint
action $a=(a_1,\ldots,a_n) \in ACT$ such that \\
$\sum_{i \in \texttt{Agt}} \tau_i(l_i(s),a^{l_i(s)\rightarrow l_i(s')},l_i(s')) > 0$ and $\sum_{s' \in S} \texttt{P}(s,s') =1$
for all $s \in S$.

\item For each pair $(i,j) \in \texttt{Agt}^2$,
$\sim_{i\rightarrow j} \subseteq S \times S$ is a social accessibility relation. $s \sim_{i\rightarrow j} s'$ is defined by the following conditions:
      %
  \begin{enumerate}
       \item $l_i(s)=l_i(s')$.
       \item $Var_i \cap Var_j \neq \emptyset$ such that $\forall x \in Var_i \cap Var_j$, we have $l^{x}_i(s)\!=\!l^{x}_j(s')$.
       \item $\forall y \in Var_j\!-\! Var_i$, we have $l^{y}_j(s)\!=\!l^{y}_j(s')$.
  \end{enumerate}
       %
  \item  $\nu : S\rightarrow 2^{AP}$ is a function valuating states with atomic propositions.
\end{itemize}
\end{definition}

\noindent Our model $\mathfrak{M_1}$ can be thought of as a labeled
state-transition system in which each transition from $s$ to $s'$
is annotated with a probability value in the matrix
$\texttt{P}$ indicating the likelihood of its occurrence
wherein the transition is assumed to take a discrete time-step.
This means that there is no notion of real time, while reasoning
about discrete time is possible through state variables keeping
track of time and counting transition steps. It is also important to
mention that every state in $\mathfrak{M_1}$ has at least one
outgoing transition to avoid deadlocks. Moreover, all
terminating/final states are modeled with a self-loop.

\vspace{0.2cm} \noindent \texttt{Computation paths}. We can unfold
the model $\mathfrak{M_1}$ into a set of paths. A path through the
model $\mathfrak{M_1}$ is a non-empty (finite or infinite) sequence
$\pi=s_0~s_1~\ldots$ of global states such that
$\texttt{P}(s_r, s_{r+1})> 0$ for all $r \geq 0$. Also,
$\pi(r)$ denotes the $(r+1)^{th}$ state of $\pi$, i.e.,
$\pi(r)=s_r$ for all $r \geq 0$.

\noindent \texttt{Probability Space}. Let $\Omega$ be a sample set
(or the set of possible outcomes of an experiment). A pair
$(\Omega,\mathrm{F})$ is said to be a \textit{sample space} if
$\mathrm{F}$ is a $\sigma$-algebra of measurable subsets of
$\Omega$, which are closed under countable union and complement
and often built from basic events called \textit{cylinders} (the
elements of $\mathrm{F}$ are called \textit{events}). A triple
$(\Omega,\mathrm{F},\mu)$ is a \textit{probability space} if
$\mu$ is a probability measure over $\mathrm{F}$, i.e., $0\leq\mu
(A)\leq 1$ for all $A\in \mathrm{F}$ such that:
%
\begin{itemize}
\item $\mu(\emptyset)=0$,
\item $\mu(\Omega) = 1$, and
\item $\mu(\bigcup_{k=1}^{\infty}A_k)=\Sigma_{k=1}^\infty \mu(A_k)$ for disjoint $A_k$.
\end{itemize}

The probability matrix $\texttt{P}$ induces a probability
space on the set of infinite paths $\Pi(s)$, which start in the
state $s$, using the cylinder construction \cite{Baier2008} as
follows. An observation of a finite path determines a basic event
(cylinder). Suppose $s = s_0$; for $\pi=s_0~s_1~\ldots~s_n$, we
define the probability measure $Prob_s \{\pi\}$ %$Prob_s^{fin}$
for the $\pi$-cylinder as follows:
%
\begin{equation}
%Prob_s^{fin}=
Prob_s \{\pi\}=
\begin{cases} 1~~~\textrm{if}~\pi~\textrm{consists~of~a~single~state}\\
{\prod_{\substack{r=0}}^{n-1} \texttt{P}(s_r,s_{r+1})~~~\textrm{otherwise.}}
 \end{cases}
\end{equation}
%
\noindent This extends to a unique measure $Prob_s$ on the set of
infinite paths $\Pi(s)$ w.r.t countable union and complement
\cite{Kwiatkowska2007}.

\subsection{Syntax of PCTLC} \label{sec:syntax-PCTLKC}

\begin{definition}[PCTLC syntax]\label{def:syntax}
Given a set of atomic propositions $AP$, the PCTLC formulae are
defined by the following BNF grammar:
%
\begin{align*}
    \varphi & ::= p~|~\neg \varphi~|~\varphi \vee \varphi~|~\mathcal{C}~|~ \mathbb{P}_{\bowtie k} (\psi)~|~\mathbb{P}_{\bowtie k}(\mathcal{C})\\
    \psi & ::=\bigcirc \varphi ~ | ~ \varphi ~U~ \varphi~|~ \varphi~ U^{\leq m} ~ \varphi \\
    \mathcal{C} & ::= C_{i\rightarrow j}\varphi ~| ~ Fu(C_{i\rightarrow j}\varphi)
\end{align*}
%
where: $p\in AP$ is an atomic proposition and $\mathbb{P}_{\bowtie
k}$ is a probabilistic operator where $\bowtie \in\{<,\leq,>,\ge\}$ and $k\in [0,1]$ is a probability bound or threshold. $m \in\mathbb{N}^+ $ is a positive integer number reflecting the maximum number of transitions needed to reach a certain state. $\varphi$ and $\psi$ are state and path formulae interpreted over the states and paths of $\mathfrak{M_1}$ respectively. The
Boolean connectives $\neg$ and $\vee$ are defined in the usual
way. Formulae $\mathcal{C}$, called social formulae, are special
state formulae in PCTLC that can express social properties using
the modal connectives $C_{i\rightarrow j}$ and $Fu(C_{i\rightarrow
j})$ standing for ``commitment'' and ``fulfillment of commitment''
respectively. $\bigcirc, U$ and $U^{\leq m}$ stand for ``next time'',
``until'' and ``bounded until'' path modal connectives respectively.

\end{definition}

\noindent The intuitive meanings of the temporal and probabilistic operators are straightforward from PCTL \cite{Jonsson1991}. $C_{i\rightarrow j}\varphi$ is read as ``agent $i$ commits towards agent $j$ that $\varphi$''. $Fu(C_{i\rightarrow j}\varphi)$ is read as ``the commitment $C_{i\rightarrow j}\varphi$ is fulfilled''. The probabilistic operator $\mathbb{P}_{\bowtie k}(\mathcal{C})$ on social formulae $\mathcal{C}$ states the degree of the commitment and the
fulfillment of the commitment: how much the agent is confident
about its commitment and fulfilling its commitment respectively.

PCTLC logic allows us to express properties like $C_{i \rightarrow j} \varphi \supset (\mathbb{P}_{\ge 0.95}(\top~U^{\leq 13}Fu(C_{i\rightarrow j}\varphi)))$ which means when a commitment about $\varphi$ is
held, then the probability that the commitment is fulfilled within
13 discrete-time steps is at least 0.95, where $\supset$ stands
for the logical implication.

\subsection{Semantics of PCTLC} \label{sec: semantics of PCTLC}

>>>>

\chapter{Coalition Formation for Autonomous Web Services}\label{sec:coalitionformationws}

In this chapter, we present our coalition model of agent-based web
services within communities\cite{SCC2013efficient}. We start by describing the general
architecture and considered parameters for web services.
Thereafter, problem modeling and formulation will be introduced in
terms of task distribution and community revenue. Web service
cooperative games in different settings will follow along with
some simulation results.

\section{Preliminaries}\label{s:preliminaries}

In this section, we discuss the parameters and preliminary
concepts that we use in the rest of the proposal.

\subsection{Architecture}

Our system consists of three main types of entities working
together:

\emph{1) Web services} are rational entities that aim to maximize
their utilities by providing high quality services to end users.
They aim to maximize their individual income by receiving enough
requests from end users. In order to increase their revenue, web
services seek for more tasks if they have the capacity and
throughput to do so. Web services can join communities to have
better efficiency by collaborating with others, to have access to
higher market share, and to have opportunity of receiving a bigger
task pool from end users. Throughout this proposal, in our
equations, we refer to web services as $ws$ and to the set of web
services hosted by a given community as $C$. To simplify the
notation, sometimes we simply write $ws$ instead of $ws \in C$ to
go through the elements $ws$ of the set $C$.

\emph{2) Master Web Services} or the community coordinators, are representatives of the
communities of web services and responsible for their management.
Communities receive requests from users and aim to host a healthy
set of web services to perform the required tasks. They seek to
maximize user satisfaction by having tasks accomplished according
to the desired QoS. In fact, higher user satisfaction will bring
more user requests and increase the market share and revenue of
the community.

\emph{3) Users} generate requests and try to find the best
available services. User satisfaction is abstracted as function of
quantity and quality of tasks accomplished by a given service.
Higher user satisfaction leads to higher trust of the community by users hence directing more requests towards that service provider.

\subsection{Web Service Parameters}\label{ws_parameters}

Web services come with different quality of service parameters.
These parameters with a short description are listed in Table
\ref{qosws}.

\begin{table}[!t]
\centering
\caption{List of web service QoS parameters.}
\begin{tabular}{|c|c||c|c|}
\hline
\textbf{Parameter} & \textbf{Definition} \\
\hline\hline
$Availability$ & Probability of being available during \\
&a time frame \\
$Reliability$ & Probability of successfully handling \\
&requests during a timeframe\\
$Successability$ & Rate of successfully handled requests \\
$Throughput$ & Average rate of handling requests \\
$Latency$ & The average latency of services\\
$Capacity$ & Amount of resources available\\
$Cost$ & Mean service fee \\
$Regulatory$ & Compliance with standards, law and rules\\
$Security$ & Quality of confidentiality \\
&and non-repudiation\\
\hline
\end{tabular}
\label{qosws}
\end{table}

We adopted a real world dataset \cite{DBLP:conf/smc/Al-MasriM09a}
which has aggregated and normalized each of these parameters to a
real value between 0 and 1. Since requests are not shared among
web services and are distributed among all of them inside a
community, each one of them comes with a given QoS denoted by
$(QoS_{ws})$. We assume that $(QoS_{ws})$ is obtained by a certain
aggregation function of the parameters considered in Table
\ref{qosws}. We use this quality output later in evaluating the
community \emph{worth} or \emph{payoff} function.

\subsection{Web Service Communities}\label{webservice-communities}

Figure 3 represents our revised architecture of web service
communities where tasks are to be distributed among the members
that are interested in forming stable coalitions. As discussed in
Chapter 2, communities are essentially virtual platforms
aggregating web services having similar and complementary
functionalities and communicate with other entities such as UDDI
registries and users using particular protocols. Web services join
communities to increase their utility by having larger market
share and task pool. Community coordinators or master web services
are responsible for community development, managing membership
requests from web services and distributing user tasks among the
community members. Community coordinators try to attract quality
web services and keep the community as stable and productive as
possible to gain better reputation and user satisfaction, which
results in having higher revenue.

\begin{figure*}[!t]
\centerline{\includegraphics[width=15cm]{Figures/archss.eps}}
\caption{Architecture of Web Service communities}
\label{fig_community}
\end{figure*}

\section{Problem Formulation and Modeling}\label{s:model}

In this section, we present  web services and community coordinator's interactions, the task distribution process and revenue models in web service communities.

\subsection{Task Distribution}

As mentioned in Section \ref{webservice-communities}, communities
are robust service providers with well established market share
and reputation. By maintaining their reputation and performance,
they attract  end users which choose them as service providers to
perform their tasks. The community master is characterized by a
request rate $(R_C)$ from users. Each web service comes with a
given QoS ($QoS_{ws}$) from which the throughput $Th_{ws}$ is
excluded. Throughput is the average rate of tasks a web service
can perform per time unit. Its exclusion from $QoS_{ws}$ allows us
to build our analysis on the particular value of $Th_{ws}$. Thus,
web services perform tasks with an average output quality of
$QoS_{ws}$ and a throughput rate of $Th_{ws}$.

The community master uses a slightly modified \emph{weighted fair queuing} method to distribute tasks among its members. The goal is to allocate incoming tasks to web services with a rate matching the throughput value of $Th_{ws}$. In \emph{weighted fair queuing} method \emph{all} the input flow is multiplexed along different paths, however in our case if the input rate $(R_C)$ of the community is more than the summation of throughput values of the web services in the community, some of the input tasks will be queued and served with delay. Thus, the amount of tasks performed by community is $\sum_{ws \in C}{(Th_{ws})}$ when $\sum_{ws}{Th_{ws}} \leq R_{C}$. However, when the input rate $(R_C)$ of the community is less than the summation of throughput values of the web services in the community,
%the community has more web services having more total throughput value than community's request rate
$(R_C)$ the \emph{weighted fair queuing} algorithm assigns a weighted task rate of $R_C \times \frac{Th_{ws}}{\sum_{ws}{Th_{ws}}}$ for each web service ($ws$) and the total rate of tasks being performed is $R_C$, the community's receiving request rate.

While distributing tasks, the community master can verify the performance, throughput and quality of service of   tasks being performed by web services. It can recognize if web services are capable of doing the amount of tasks they advertised. If for any reason there is a decline in quality metrics or throughput, the  community master will announce the new parameters and community masters and members can consider those values as benchmark for future performance calculations but also to penalize them.
In this way, players have incentive to reveal their real capabilities to profit best from the community and to avoid being penalized. In addition, the system should be dynamic enough to detect and react to web services quality metrics variation as over time web service metrics may  degrade or improve, a change that the community should adjust to.
% Therefore its easy for the system to encourage players to be in some sense incentive compatible in the way that they would profit best by truthfully revealing their capabilities. Also it is important to be dynamic enough to consider web services which may have their quality metrics degraded or even improved over time for any reason and be able to adjust the community with new parameters.

\subsection{Community Revenue}

The communities and web services earn revenue by performing tasks.
The total gain is function of quality ($QoS_{ws}$) and throughput
($Th_{ws}$) of tasks being performed. As mentioned in section
\ref{ws_parameters}, $QoS_{ws}$ is obtained by a certain
aggregation function of the parameters considered in Table
\ref{qosws}. We have adopted a linear equal weight average over
all QoS parameters listed in Table \ref{qosws} excluding the
$Throughput$ and $Cost$ parameters. A community has the option to
weigh specific QoS parameters depending on the expectations of
their clients.

The maximum potential output of a community $(PO(C))$  is an aggregation of number of tasks, times their quality, for each web service member of the community:

\begin{equation}
PO(C) = \sum_{ws \in C}{(T_{ws} \times QoS_{ws})}
\end{equation}

If the summation of throughput values ($Th_{ws}$) of community members exceeds the input task rate of the community ($R_C$) the community cannot perform at its maximum potential. It denotes the case when the community has more web services than it needs to perform the input task load. The actual output has to be normalized to the amount of tasks being performed.

\begin{equation}\label{out_c}
Out(C) = \left\{
  \begin{array}{l l}
    PO(C) & \quad \text{if $\sum_{ws}{Th_{ws}} \leq R_{C}$}\\
    PO(C) \times \frac{R_{C}}{\sum_{ws}{Th_{ws}}} & \quad \text{if $\sum_{ws}{Th_{ws}} > R_{C}$}
  \end{array} \right.
\end{equation}

The revenue function of the web service community is a linear function of $Out(C)$ with a positive constant multiplier.

\subsection{Case Study}

In this section, we analyze three numerical examples and discuss the motivation of web services and community interactions and the strategies they can adopt and the revenue they can earn adopting these different strategies.

%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Case Study: Example 1}
\label{example_1}
\centering
\begin{tabular}{c c c c}
\hline
$WS$ & $QoS_{ws}$ & $Th_{ws}$ & $Th_{ws} \times QoS_{ws}$\\
\hline
1 & 0.8 & 4 & 3.2\\
2 & 0.8 & 5 & 4.0\\
3 & 0.8 & 3 & 2.4\\
\hline
\end{tabular}
\end{table}

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
% \caption{Three web services}
\label{example_1_2}
\centering
\begin{tabular}{c c || c c}
\hline
Community & Worth & Community & Worth\\
\hline
$\left\{1\right\}$ & 3.2 & $\left\{1,2\right\}$ & 7.2\\
$\left\{2\right\}$ & 4.0 & $\left\{1,3\right\}$ & 5.6\\
$\left\{3\right\}$ & 2.4 & $\left\{2,3\right\}$ & 6.4\\
$\left\{1,2,3\right\}$ & 8.0\\
\hline
Community $R_C$: 10\\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 2 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Case Study: Example 2}
\label{example_2}
\centering
\begin{tabular}{c c c c}
\hline
$WS$ & $QoS_{ws}$ & $Th_{ws}$ & $Th_{ws} \times QoS_{ws}$\\
\hline
1 & 0.8 & 5 & 4.0\\
2 & 0.7 & 6 & 4.2\\
3 & 0.7 & 4 & 2.8\\
\hline
\end{tabular}
\end{table}

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
% \caption{Three web services}
\label{example_2_2}
\centering
\begin{tabular}{c c || c c}
\hline
Community & Worth & Community & Worth\\
\hline
$\left\{1\right\}$ & 4.0 & $\left\{1,2\right\}$ & 7.4\\
$\left\{2\right\}$ & 4.2 & $\left\{1,3\right\}$ & 6.8\\
$\left\{3\right\}$ & 2.8 & $\left\{2,3\right\}$ & 7.0\\
$\left\{1,2,3\right\}$ & 7.3\\
\hline
Community $R_C$: 10\\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 2 %%%%%%%%%%%%%%, %%%%%%%%%%%%%%%%%%%%%%

In the first example,  we present the case of a community with
$R_C =10 $, and three web services, each having different
$QoS_{ws}$ and $Th_{ws}$ values as listed in Table
\ref{example_1}. The worth of a community is calculated based on
$Out(C)$ equation (\ref{out_c}) which is the amount of output
being generated by the community. The first table  lists the web
services with their aggregated $QoS_{ws}$ parameters, their task
input rate while working alone, and also their  throughput value
$Th_{ws}$. The second table shows all the possible communities and
their respective worth. The obtained values suggest that
communities having more web services have better gain and output.
However each community needs to  distribute the gain between web
services. Sometimes it is impossible to share the gain between all
web services in a way that no subset of them would individually
gain more if they form their own group. In this example, the value
community of ${ws_1}$ and ${ws_2}$ is 7.2, With ${ws_3}$ joining
the community the worth increases to 8.0. However there is no way
to distribute the value among web services to have  ${ws_1}$ and
${ws_2}$  earning 7.2, and ${ws_3}$ earning at least 2.4, the gain
they could earn before joining the community. This fact makes the
group unstable. In the second  example, shown in Table
\ref{example_2}, we even have situations where a web service
(${ws_3}$) joining a community ($\left\{ws_1,ws_2\right\}$)
decreases the value of community. The reason is, the community is
already full and all tasks are almost being distributed and new
community with bad quality can degrade the average quality of
tasks being done by the community. In both examples, the request
of joining of web service ${ws_3}$ should be rejected by the
community.

%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Case Study: Example 3}
\label{example_3}
\centering
\begin{tabular}{c c c c}
\hline
$WS$ & $QoS_{ws}$ & $Th_{ws}$ & $\text{\emph{Input Task Rate}}$\\
\hline
1 & 0.8 & 10 & 5\\
2 & 0.8 & 20 & 5\\
3 & 0.8 & 30 & 5\\
\hline
\end{tabular}
\end{table}

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
% \caption{Three web services}
\label{example_3_2}
\centering
\begin{tabular}{c c || c c}
\hline
Community & Worth & Community & Worth\\
\hline
$\left\{C_{ms_1}\right\}$ & 0 & $\left\{C_{ms_2}\right\}$ & 0\\
$\left\{C_{ms_1}, ws_1\right\}$ & 8 & $\left\{C_{ms_2}, ws_1\right\}$ & 8\\
$\left\{C_{ms_1}, ws_2\right\}$ & 16 & $\left\{C_{ms_2}, ws_2\right\}$ & 16\\
$\left\{C_{ms_1}, ws_3\right\}$ & 16 & $\left\{C_{ms_2}, ws_3\right\}$ & 24\\
$\left\{C_{ms_1}, ws_1, ws_2\right\}$ & 16 & $\left\{C_{ms_2}, ws_1, ws_2\right\}$ & 24\\
$\left\{C_{ms_1}, ws_1, ws_3\right\}$ & 16 & $\left\{C_{ms_2}, ws_1, ws_3\right\}$ & 32\\
$\left\{C_{ms_1}, ws_2, ws_3\right\}$ & 16 & $\left\{C_{ms_2}, ws_2, ws_3\right\}$ & 32\\
$\left\{C_{ms_1}, ws_1, ws_2, ws_3\right\}$ & 16 & $\left\{C_{ms_2}, ws_1, ws_2, ws_3\right\}$ & 32\\
$\left\{C_{ms_1}, C_{ms_2}, ...\right\}$ & 0 & $\left\{ws_1\right\}$ & 6.8\\
$\left\{ws_2\right\}$ & 4.2 & $\left\{ws_3\right\}$ & 6.8\\
\hline
Community $R_{C_1}$: 20 \\ Community $R_{C_2}$: 40\\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%% EXAMPLE 3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In Example 3, we consider the case of having different communities with different market share, ${R_C}$ values. Web services also have a small share of market independently, providing them with a small task pull. In these kind of scenarios, the solution considers individual maximization of payoff and also the total worth of all communities which represents the \emph{social welfare}. In this example the most efficient partition of web services is earned by having two coalitions of $\left\{C_{master_1}, ws_2\right\}$ and $\left\{C_{master_2}, ws_1, ws_3\right\}$, which yields a total value of $32 + 16 = 48$. In these types of scenarios, the goal is to reach stability, adopting a distributed approach where all players have the power of choice on the decision of whether or not they join a coalition. The communities usually start the game having some established members, encountering new web services, the communities may exchange web services and new web services would join them having at least one player gaining utility, without hurting any other participant. In this example if we initially having two coalitions of $\left\{C_{master_1}, ws_2\right\}$ and $\left\{C_{master_2}, ws_1\right\}$ and a ${ws_3}$ as new web service, ${ws_3}$ joining ${C_{master_1}}$ would hurt at least itself or $ws_2$, however ${ws_3}$ joining ${C_{master_2}}$ would not hurt any participants and ${ws_3}$ would earn more within the community and the community will have enough web services performing the incoming tasks from users.

The first two examples illustrate the fact that a community cannot simply increase its revenue by adding more web services. The web services and even community owners are autonomous agents and would deviate and be displeased about the community if new members cause a drop in their profit. The job of the community master is to attract as many quality web services it can and keep them satisfied; hence the group stability is guaranteed.
The third example highlights another type of problem we would like to address, which is how to form best possible groups of communities, and allocate web services among communities in a way which would maximize payoff for of our agents and members already residing in the communities.
In next section, we provide collaborative game theory based algorithms for our autonomous agents, to tackle these problems and find applicable and efficient strategies for communities and web services to maximize their profit.

 \section{Web Service Cooperative Games}\label{s:game_solution}

In this section, we present different web service community models
and focus on the problem of how both web services and community
masters as rational entities would adopt strategies to maximize
their payoff.

\subsection {Web Services and One Community}
